% !TEX root = ../main.tex
\chapter*{Acknowledgement / Ringraziamenti}

Given the length of this work, it is no wonder even the acknowledgements are going to be somewhat long --- for those who don't know me: I am known for being long-winded. 

First and foremost, I would like to express my deepest gratitude to my supervisor, Alice Campani, who encouraged me to explore every idea that came to mind while also ensuring that I remained focused on the main tasks and discussing any results I obtained. I strongly believe that I am a better critical thinker now, and that is mostly thanks to you. 
I would also like to express my gratitude to Lea Di Noto and Sergio Di Domizio for their insightful comments on my work and for helping me to identify the right questions to ask. 

Furthermore, I would like to express my gratitude to everyone who has played a key role in my daily activities over the past year: Ricardo Triozzi, María Artero Pons, Christian Farnese, and all the wonderful people in Padua, who were extremely kind and willing to collaborate. 
Finally, I would like to express my gratitude to all the wonderful people from the ICARUS collaboration whom I had the pleasure of meeting in person at Fermilab. They convinced me that pursuing this thesis was the right decision. 


Non posso però dimenticare chi, da casa, ha fatto, anche indirettamente, la sua parte. Innanzitutto Fra, con il quale ho condiviso gli ultimi 10 anni di studi e pomeriggi interi di laboratorio. Grazie. Poi Matti, Andre e Flo, grazie per i pranzi, le pause caffè e per aver ripetuto il programma di ogni esame. 
Grazie a tutti i ``residenti'' di fisica che hanno con me condiviso questi ultimi cinque anni.

Grazie infine a Giulia, Annalea, Lorenzo e Ricardo, per esserci stati anche da distante.

Grazie alla mia famiglia, a mia sorella, che mi ha sostenuto sempre, anche quando non capivano cosa facessi, anche quando tornavo stanco e non vi dicevo nulla: grazie. Ai nonni Pupa e Giovanni che ci sono sempre stati: grazie; Anche se non è qua, so che anche nonna Mirella è felice per la mia laurea; farà strano solo che non potrò raccontarglielo. Questa tesi è dedicata a tutti voi, che mi avete sostenuto. 

\begin{flushright}
    \emph{Mattia}\par
    16 ottobre 2025
\end{flushright}